---
title: "Econ PSET 1"
author: "Kevin Li"
date: "1/11/2020"
output:
  pdf_document: 
    keep_tex: yes
  html_document: default
---

```{r setup, include=FALSE}
# Load some commonly used packages.
# If you get a "package not found" error, 
# then follow the instructions for installing packages at 
# http://statistics.uchicago.edu/~collins/Rinstall/
library(MASS)
library(gridExtra)
library(tidyverse)
library(mosaic)
library(broom)
library(dplyr)
library(prob)
library(reshape2)
library(R.utils)
library(plyr)
library(nloptr)
library(rootSolve)
library(openxlsx)
library(tidyr)
library(Hmisc)
# Set numerical output display parameters
options(width=70, digits=4, scipen=8)
# Set R output size a bit smaller than default
knitr::opts_chunk$set(size='small', prompt=FALSE, comment="")
# set plot theme to black and white
ggplot2::theme_set(ggplot2::theme_bw())
ggplot2::theme_update(panel.grid.major = ggplot2::element_line(colour = "grey75"))
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```


Read all the data. Remember to comment out the file path when doing it yourself 
```{r}
setwd("~/Downloads/2020 Spring Files/Econ /Datasets/HortasuProject")
alcdrugs = read.csv("alcdrugs.csv", header = TRUE)
count_ids <- read.delim("For_County_IDs.txt", sep = "\t")
mls_df = read.csv("mls.csv", header = TRUE)


```

## OLS Estimation


This block below cleans up the data and sets up the three key dataframes
```{r}
#cleans alcdrug dataframe
get_cleaned_alcdrugs = function(raw_alcdrug_df){
  relevant = raw_alcdrug_df[raw_alcdrug_df$Drug.Alcohol.Induced.Code != "O", ]
  relevant$ID = paste(relevant$County.Code, relevant$Year, sep="_")
  wide_df = dcast(relevant, ID + Population ~Drug.Alcohol.Induced.Code, value.var="Deaths")
  return(wide_df)
}


#prepares mls_df dataframe
process_mls = function(mls_df)
{
  mls_df
  mls_df$ID <- paste(mls_df$State_county.FIPS, mls_df$Year, sep = "_")
  return(mls_df)
}


#get shell dataframe, all county years
get_shell = function(count_ids)
{
  county_ids <- count_ids$County.Code
  years <- c(1999:2012)
  #I realized I needed to get a complete set of county_ids
  all_county_years <- expand.grid(years, county_ids)
  colnames(all_county_years) <- c("Year", "County_code")
  all_county_years$ID <- paste(all_county_years$County_code, 
                               all_county_years$Year, sep = "_")
  
  all_county_years
}


cleaned_alcdrugs = get_cleaned_alcdrugs(alcdrugs)
mls_df = process_mls(mls_df)
all_county_years = get_shell(count_ids)
```



This block merges the three dataframes and runs the basic analysis:

```{r}
#merges three files together and adds some minor changes
create_analysis_dataframe = function(all_county_years, mls_df, cleaned_alcdrugs)
{
  complete_mls = merge(x = all_county_years, y= mls_df, by = "ID", all.x = TRUE)
  analysis_df = merge(x = complete_mls, y = cleaned_alcdrugs, by = "ID", all.x = TRUE)
  analysis_df$total_deaths = analysis_df$A + analysis_df$D
  
  analysis_df$Total[is.na(analysis_df$Total)] <- 0 #fills in 0 layoffs for missing data in BLS
  analysis_df$layoff_perc = analysis_df$Total / analysis_df$Population #adds a proportion
  
  return(analysis_df)
}

analysis_df = create_analysis_dataframe(all_county_years, mls_df, cleaned_alcdrugs)
analysis_df
```


Here, you can try to run all your regressions on the most basic type (no lag variables)

```{r}
first_regression = lm(data=analysis_df, formula = total_deaths ~ Total + factor(Year.x) + factor(County_code))
summary(first_regression)
```


Next, we're going to consider regressions with lag variables. The functin below helps you create a mls dataframe with lag variables

```{r}
#merges three files together and adds a lag variable of one year, 
#abandoned: even for one year this is impractically slow
slow_merge_with_lag = function(all_county_years, mls_df, cleaned_alcdrugs, lag_years = 1)
{
  complete_mls = merge(x = all_county_years, y= mls_df, by = "ID", all.x = TRUE)
  complete_mls$Total[is.na(complete_mls$Total)] <- 0 #fills in 0 layoffs for missing data in BLS
  
  lag_dataset = data.frame()
  #IMPORTANT: This function assumes the dataframe is already sorted by time
  for (county_id in unique(complete_mls$County_code)){
    county_df = complete_mls[complete_mls$County_code == county_id, ]
    county_df = county_df[!is.na(county_df$ID), ]
    county_df$lag1 <- Lag(county_df$Total, +1)
    county_df = county_df[2:nrow(county_df), ]
    
    lag_dataset = rbind(lag_dataset, county_df)
  }
  
  return(lag_dataset)
}
```


```{r}

# Improved Version of Merged with Lag that isn't stupid Slow
#
# This only works if you're sorting by County ID then Year
# My ID's differ from the main.r file for this reason
# Do not put lag >= 11
merge_with_lag = function(all_county_years, mls_df, cleaned_alcdrugs, lag_years = 1)
{
  relevant_years = c(1999, 2000, 2001, 2002, 2003, 2004, 2005, 
                     2006, 2007, 2008, 2009, 2010, 2011, 2012)
  
  complete_mls = merge(x = all_county_years, y= mls_df, by = "ID", all.x = TRUE)
  complete_mls$Total[is.na(complete_mls$Total)] <- 0 #fills in 0 layoffs for missing data in BLS
  
  for (lag_year in 1:lag_years)
  {
    name = paste("lag", lag_year, sep="_")
    complete_mls[ , name]= Lag(complete_mls$Total, +lag_years)
    relevant_years = relevant_years[2: length(relevant_years)]
  }
  
  complete_mls = complete_mls[complete_mls$Year.x %in% relevant_years,  ]
  return(complete_mls)
}

test = merge_with_lag(all_county_years, mls_df, cleaned_alcdrugs, lag_years = 3)
test
```